---
title: "Modeling Jobs"
format: html
warning: false
message: false
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(stringr)
```


# Objective:

End goal is to have one data frame that looks something like this:

| county  | year | technology | new_capacity | total_capacity | occupation   | type    | n_jobs | ambition 
|---------|------|------------|--------------|----------------|--------------|---------|--------|
|   SLO   | 2026 | Utility PV | 1            | 1              | construction |direct   | 5      | high
| ...     | ...  |  ...       | ...          | ...            | ...          |  ...    | ...    | low 
| Ventura | 2026 | OSW        | 1            | 1              | op and maint |indirect | 100    |

etc.

I imagine this coming together (concat) from many different dataframes: one for each county and technology

# PV model

Start with 1 piece and see if we can build this into a function from there

## Santa Barbara Utility Solar

Values from Eleanor's Tech. Scenarios sheet:

- Ambition: 722 MW by 2045
- Starting: 111 MW in 2025

First, use compounding interest to calculate yearly capacities:

```{r}
# Initial values
initial_capacity <- 111
final_capacity <- 722
years <- 20

# Calculate the annual growth rate
growth_rate <- (final_capacity / initial_capacity)^(1 / years) - 1

# Create variables to store the results
year <- 2025:2045
capacity <- numeric(length(year))
new_capacity <- numeric(length(year))

# Calculate the total capacity for each year
for (i in 1:length(year)) {
  capacity[i] <- initial_capacity * (1 + growth_rate)^(year[i] - 2025)
  if (i == 1) {
      new_capacity[i] <- capacity[i] - initial_capacity
  } else {
      new_capacity[i] <- capacity[i] - capacity[i-1]
  }
}

# Create a data frame with the results
sb_utility_pv <- data.frame(county = "SB", 
                             year = year, 
                             technology = "utility solar",
                             new_capacity = new_capacity,
                             total_capacity = round(capacity, 2))

# Display the table
print(sb_utility_pv)
```

Now we just need to add number of jobs! Let's just say for testing purposes that utility solar requires:
- 5 FTE construction jobs / MW
- construction project lasts 3 months
- 10 annual O&M jobs / MW
- lifetime of project is 30 years (beyond the scope of our model)  


Starting with O&M jobs:
```{r}
# O&M direct jobs ---- 
# Assign variables
om_direct <- 10

# Add columns occupation, type, and n_jobs
sb_utility_pv_om_direct <- sb_utility_pv |>
    mutate(occupation = "O&M",
           type = "direct",
           n_jobs = total_capacity * om_direct)

# O&M indirect jobs ---- 
# Assign variables
om_indirect <- 5

# Add columns occupation, type, and n_jobs
sb_utility_pv_om_indirect <- sb_utility_pv |>
    mutate(occupation = "O&M",
           type = "indirect",
           n_jobs = total_capacity * om_indirect)

# O&M induced ----
# Assign variables
om_induced <- 2

# Add columns occupation, type, and n_jobs
sb_utility_pv_om_induced <- sb_utility_pv |>
    mutate(occupation = "O&M",
           type = "induced",
           n_jobs = total_capacity * om_induced)

# Stack 3 together for total SB PV O&M jobs
sb_utility_pv_om <- rbind(sb_utility_pv_om_direct, sb_utility_pv_om_indirect, sb_utility_pv_om_induced)

print(sb_utility_pv_om)

```

Nice. Let's build this into a function, starting with a function for O&M jobs:
```{r}
# Function to calculate capacity and jobs for PV O&M
pv_om_capacity_and_jobs <- function(county, years, technology, initial_capacity, final_capacity, direct_jobs, indirect_jobs, induced_jobs) {
    
  # Calculate the annual growth rate
  growth_rate <- (final_capacity / initial_capacity)^(1 / years) - 1
  
  # Create variables to store the results
  year <- 2025:2045
  capacity <- numeric(length(year))
  new_capacity <- numeric(length(year))
  
  # Calculate the total capacity for each year
  for (i in 1:length(year)) {
    capacity[i] <- initial_capacity * (1 + growth_rate)^(year[i] - 2025)
    if (i == 1) {
      new_capacity[i] <- capacity[i] - initial_capacity
    } else {
      new_capacity[i] <- capacity[i] - capacity[i-1]
    }
  }
  
  # Create a data frame with the results
  df <- data.frame(county = county, 
                   year = year, 
                   technology = technology,
                   new_capacity = new_capacity,
                   total_capacity = round(capacity, 2))
  
  # Direct jobs
  df_direct <- df %>%
    mutate(occupation = "O&M", 
           type = "direct", 
           n_jobs = total_capacity * direct_jobs)
  
  # Indirect jobs
  df_indirect <- df %>%
    mutate(occupation = "O&M",
           type = "indirect", 
           n_jobs = total_capacity * indirect_jobs)
  
  # Induced jobs
  df_induced <- df %>%
    mutate(occupation = "O&M",
           type = "induced", 
           n_jobs = total_capacity * induced_jobs)
  
  # Stack them together for total jobs
  df_final <- rbind(df_direct, df_indirect, df_induced)
  
  return(df_final)
}

```


And then a function for Construction jobs
```{r}
# Function to calculate capacity and jobs for PV Construction
pv_const_capacity_and_jobs <- function(county, years, technology, initial_capacity, final_capacity, direct_jobs, indirect_jobs, induced_jobs) {
    
  # Calculate the annual growth rate
  growth_rate <- (final_capacity / initial_capacity)^(1 / years) - 1
  
  # Create variables to store the results
  year <- 2025:2045
  capacity <- numeric(length(year))
  new_capacity <- numeric(length(year))
  
  
  # total_capacity[i] <- initial_capacity * (1 + growth_rate)^(year[i] - 2030)
  # Calculate the total capacity for each year
  for (i in 1:length(year)) {
    capacity[i] <- initial_capacity * (1 + growth_rate)^(year[i] - 2025)
    if (i == 1) {
      new_capacity[i] <- capacity[i] - initial_capacity
    } else {
      new_capacity[i] <- capacity[i] - capacity[i-1]
    }
  }
  
  # Create a data frame with the results
  df <- data.frame(county = county, 
                   year = year, 
                   technology = technology,
                   new_capacity = new_capacity,
                   total_capacity = round(capacity, 2))
  
  # Direct jobs
  df_direct <- df %>%
    mutate(occupation = "Construction", 
           type = "direct", 
           n_jobs = new_capacity * direct_jobs)   # Assuming that construction jobs only last the year, #jobs/mw year will multiply by the new capacity
  
  # Indirect jobs
  df_indirect <- df %>%
    mutate(occupation = "Construction",
           type = "indirect", 
           n_jobs = new_capacity * indirect_jobs)
  
  # Induced jobs
  df_induced <- df %>%
    mutate(occupation = "Construction",
           type = "induced", 
           n_jobs = new_capacity * induced_jobs)
  
  # Stack them together for total jobs
  df_final <- rbind(df_direct, df_indirect, df_induced)
  
  return(df_final)
}
```

Let's use the functions for each county and occupation
```{r}
# Utility PV
sb_utility_pv_om <- pv_om_capacity_and_jobs(
  county = "SB", 
  technology = "Utility PV", 
  years = 20,
  initial_capacity = 110.8566193, 
  final_capacity = 722.0797068, 
  direct_jobs = 0.2, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

sb_utility_pv_const <- pv_const_capacity_and_jobs(
  county = "SB",
  years = 20,
  technology = "Utility PV",
  initial_capacity = 110.8566193,
  final_capacity = 722.0797068,
  direct_jobs = 1.6,
  indirect_jobs = 0.6,
  induced_jobs = 0.4
)

slo_utility_pv_om <- pv_om_capacity_and_jobs(
  county = "SLO", 
  years = 20,
  technology = "Utility PV", 
  initial_capacity = 1615.819209, 
  final_capacity = 10524.85876, 
  direct_jobs = 0.2, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

slo_utility_pv_const <- pv_const_capacity_and_jobs(
  county = "SLO",
  years = 20,
  technology = "Utility PV",
  initial_capacity = 1615.819209,
  final_capacity = 10524.85876,
  direct_jobs = 1.6,
  indirect_jobs = 0.7,
  induced_jobs = 0.3
)

ventura_utility_pv_om <- pv_om_capacity_and_jobs(
  county = "V", 
  years = 20,
  technology = "Utility PV", 
  initial_capacity = 6.71858299, 
  final_capacity = 43.76240647, 
  direct_jobs = 0.2, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

ventura_utility_pv_const <- pv_const_capacity_and_jobs(
  county = "V",
  years = 20,
  technology = "Utility PV",
  initial_capacity = 6.71858299, 
  final_capacity = 43.76240647, 
  direct_jobs = 1.6,
  indirect_jobs = 0.6,
  induced_jobs = 0.4
)


# Rooftop PV
sb_roof_pv_om <- pv_om_capacity_and_jobs(
  county = "SB", 
  years = 20,
  technology = "Rooftop PV", 
  initial_capacity = 242.0159119, 
  final_capacity = 1293.941196, 
  direct_jobs = 0.3, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

sb_roof_pv_const <- pv_const_capacity_and_jobs(
  county = "SB",
  years = 20,
  technology = "Rooftop PV",
  initial_capacity = 242.0159119,
  final_capacity = 1293.941196,
  direct_jobs = 5.9,
  indirect_jobs = 4.7,
  induced_jobs = 2.5
)

slo_roof_pv_om <- pv_om_capacity_and_jobs(
  county = "SLO", 
  years = 20,
  technology = "Rooftop PV", 
  initial_capacity = 344.8405982, 
  final_capacity = 1843.694708, 
  direct_jobs = 0.3, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

slo_roof_pv_const <- pv_const_capacity_and_jobs(
  county = "SLO",
  years = 20,
  technology = "Rooftop PV",
  initial_capacity = 344.8405982,
  final_capacity = 1843.694708,
  direct_jobs = 6.2,
  indirect_jobs = 5.4,
  induced_jobs = 2.3
)

ventura_roof_pv_om <- pv_om_capacity_and_jobs(
  county = "V", 
  years = 20,
  technology = "Rooftop PV", 
  initial_capacity = 424.1984954, 
  final_capacity = 3026.377541, 
  direct_jobs = 0.3, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

ventura_roof_pv_const <- pv_const_capacity_and_jobs(
  county = "V",
  years = 20,
  technology = "Rooftop PV",
  initial_capacity = 424.1984954,
  final_capacity = 3026.377541,
  direct_jobs = 6.2,
  indirect_jobs = 4.7,
  induced_jobs = 2.5
)

# Commercial PV
sb_comm_pv_om <- pv_om_capacity_and_jobs(
  county = "SB", 
  years = 20,
  technology = "Commercial PV", 
  initial_capacity = 242.0159119, 
  final_capacity = 1293.941196, 
  direct_jobs = 0.2, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

sb_comm_pv_const <- pv_const_capacity_and_jobs(
  county = "SB",
  years = 20,
  technology = "Commercial PV",
  initial_capacity = 242.0159119,
  final_capacity = 1293.941196,
  direct_jobs = 3.1,
  indirect_jobs = 2.9,
  induced_jobs = 1.5
)

slo_comm_pv_om <- pv_om_capacity_and_jobs(
  county = "SLO", 
  years = 20,
  technology = "Commercial PV", 
  initial_capacity = 344.8405982, 
  final_capacity = 1843.694708, 
  direct_jobs = 0.2, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

slo_comm_pv_const <- pv_const_capacity_and_jobs(
  county = "SLO",
  years = 20,
  technology = "Commercial PV",
  initial_capacity = 344.8405982,
  final_capacity = 1843.694708,
  direct_jobs = 3.3,
  indirect_jobs = 3.3,
  induced_jobs = 1.3
)

ventura_comm_pv_om <- pv_om_capacity_and_jobs(
  county = "V", 
  years = 20,
  technology = "Commercial PV", 
  initial_capacity = 566.0469908, 
  final_capacity = 3026.377541, 
  direct_jobs = 0.2, 
  indirect_jobs = 0, 
  induced_jobs = 0
)

ventura_comm_pv_const <- pv_const_capacity_and_jobs(
  county = "V",
  years = 20,
  technology = "Commercial PV",
  initial_capacity = 566.0469908,
  final_capacity = 3026.377541,
  direct_jobs = 3.3,
  indirect_jobs = 2.9,
  induced_jobs = 1.4
)


# Combine all counties and technologies --> for our purposes, we'll keep rooftop as residential, and add in commercial
pv_all <- rbind(sb_utility_pv_om, sb_utility_pv_const, slo_utility_pv_om, slo_utility_pv_const, ventura_utility_pv_om, ventura_utility_pv_const,
                sb_roof_pv_om, sb_roof_pv_const, slo_roof_pv_om, slo_roof_pv_const, ventura_roof_pv_om, ventura_roof_pv_const, sb_comm_pv_om,
                sb_comm_pv_const, slo_comm_pv_om, slo_comm_pv_const, ventura_comm_pv_om, ventura_comm_pv_const)
```

NEED TO ADD IN ABMITION SCENARIOS! Make a new column with ambition = high or low.

Base plot to refer back to, using Ventura utility as an example

```{r, fig.height= 6, fig.width= 8}
pv_v_plot <- pv_all |>
    filter(type == "direct" &
           county == "V" &
           technology == "Utility PV") 


pv_all_plot1 <- ggplot(pv_v_plot) +
    theme_minimal() +
    geom_bar(aes(x = year, y = n_jobs, fill = occupation), size = 1.25, position = "stack", stat="identity") +
    scale_fill_manual(values = c("Construction" = '#4a4e69', "O&M" = '#9a8c98'))+ 
    labs(x = NULL,
         y= "FTE jobs") +
    scale_y_continuous(labels = scales::label_comma()) +
    scale_fill_manual(labels = c("Construction Jobs", "Operations & Maintenance Jobs"),
                      values = c("#4a4e69", "#9a8c98")) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, margin = margin(5,20,0,10)),
        axis.text = element_text(size = 20), 
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.position = "bottom",
        plot.background = element_rect(fill = "#EFEFEF"),
        plot.title = element_blank(),
        panel.grid = element_line(color = "grey85")
    )


pv_all_plot1


ggsave('plots/pv_base_plot.png', pv_all_plot1, width = 8, height = 6)

```

Plot of pv capacities

```{r}
#pv_cap_plot 

pv_capacity_growth_plot_V <- ggplot(ventura_utility_pv_const) +
    theme_minimal() +
    geom_line(aes(x = year, y = new_capacity), color = "#9a8c98", lwd = 2) +
    geom_line(aes(x = year, y = total_capacity), color = "#4a4e69", lwd = 2) +
    theme(plot.background = element_rect(color = "black", fill = NA, size = 1),
          legend.text = element_blank(),
            axis.text.x = element_text(size = 16),  # Rotate x-axis labels
    axis.text.y = element_text(size = 16), 
          legend.position = 'none') +
    labs(x = NULL,
         y= "Capacity (GW)") +
    scale_y_continuous(labels = scales::label_comma()) +
    scale_fill_manual(labels = c("New capacity", "Total capacity"),
                      values = c("#4a4e69", "#9a8c98")) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, margin = margin(5,20,0,10)),
        axis.text = element_text(size = 20), 
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.position = "bottom",
        plot.background = element_rect(fill = "#EFEFEF"),
        plot.title = element_blank(),
        panel.grid = element_line(color = "grey85")
    )

ggsave('plots/pv_capacity_growth_plot_V.png', pv_capacity_growth_plot_V, width = 8, height = 6)
pv_capacity_growth_plot_V
```

Plot of osw capacities

```{r}
total_cap_line <- osw |>
    filter(year %in% c(2030:2045))
annual_cap_line <- osw |>
    filter(year %in% c(2026:2041))

osw_capacity_growth_plot_faculty_review <- ggplot() +
    geom_line(data = total_cap_line, 
              aes(x = year, y = total_capacity),
              color = "#9a8c98",
              lwd = 2) +
    geom_line(data = annual_cap_line, 
              aes(x = year, y = new_construction_capacity), 
              color = "#4a4e69",
              lwd = 2) +
    scale_x_continuous(limits = c(2025,2045)) +
    scale_y_continuous(limits = c(0,15),
                       breaks = c(0,3,5,10,15)) +
    labs(y = "Capacity (GW)") +
    theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, margin = margin(5,20,0,10)),
        axis.text = element_text(size = 20), 
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.position = "bottom",
        plot.background = element_rect(fill = "#EFEFEF"),
        plot.title = element_blank(),
        panel.grid = element_line(color = "grey85")
    )

ggsave("plots/osw_capacity_growth_plot_faculty_review.png",
       osw_capacity_growth_plot_faculty_review,
       width = 10,
       height = 7)
```


Utility PV in Ventura

```{r, fig.height= 7, fig.width= 10}
# 2nd plot attempt
pv_u_v_plot <- pv_all |>
    filter(type == "direct" &
           county == "V" &
           technology == "Utility PV") 


pv_u_v_plot <- ggplot(pv_u_v_plot) +
    theme_minimal() +
    geom_bar(aes(x = year, y = n_jobs, fill = occupation), size = 1.25, position = "stack", stat="identity")+
    scale_fill_manual(values = c("Construction" = '#4a4e69', "O&M" = '#9a8c98')) + 
    labs(x = NULL,
         y= "FTE jobs",
         title = "Projected direct jobs in Ventura county from Utility PV development") +
    scale_y_continuous(labels = scales::label_comma()) +
    scale_fill_manual(labels = c("Construction Jobs", "Operations & Maintenance Jobs"),
                      values = c("#4a4e69", "#9a8c98")) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, margin = margin(5,20,0,10)),
        axis.text = element_text(size = 20), 
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.position = "bottom",
        plot.background = element_rect(fill = "#EFEFEF"),
        plot.title = element_blank(),
        panel.grid = element_line(color = "grey85")
    )


pv_u_v_plot


ggsave('plots/pv_u_v_plot.png', pv_u_v_plot, width = 10, height = 7)
```

Utility PV in Santa Barbara

```{r, fig.height= 6, fig.width= 8}
pv_u_sb_plot <- pv_all |>
    filter(type == "direct" &
           county == "SB" &
           technology == "Utility PV") 


pv_u_sb_plot1 <- ggplot(pv_u_sb_plot) +
    theme_minimal() +
    geom_bar(aes(x = year, y = n_jobs, fill = occupation), size = 1.25, position = "stack", stat="identity")+
    scale_fill_manual(values = c("Construction" = '#4a4e69', "O&M" = '#9a8c98'))+ 
    #scale_x_continuous(expand = c(0, 0), breaks = scales::pretty_breaks()) +  # Remove space on x-axis
 # scale_y_continuous(expand = c(0, 0), breaks = scales::pretty_breaks()) +   # Remove space on y-axis 
    theme(plot.background = element_rect(color = "black", fill = NA, size = 1),
          #legend.text = element_blank(),
            axis.text.x = element_text(size = 16),  # Rotate x-axis labels
    axis.text.y = element_text(size = 16))+
    labs(x = NULL,
         y= "FTE jobs",
         title = "Projected direct jobs in Santa Barbara county from Utility PV development") +
    scale_y_continuous(labels = scales::label_comma()) +
    scale_fill_manual(labels = c("Construction Jobs", "Operations & Maintenance Jobs"),
                      values = c("#4a4e69", "#9a8c98")) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, margin = margin(5,20,0,10)),
        axis.text = element_text(size = 20), 
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.position = "bottom",
        plot.background = element_rect(fill = "#EFEFEF"),
        plot.title = element_blank(),
        panel.grid = element_line(color = "grey85")
    )



pv_u_sb_plot1


ggsave('plots/pv_u_sb_plot.png', pv_all_plot1, width = 8, height = 6)
```

<!-- Utility PV in San Luis Obispo -->

```{r, fig.height= 6, fig.width= 8}
pv_u_slo_plot <- pv_all |>
    filter(type == "direct" &
           county == "SLO" &
           technology == "Utility PV") 


pv_u_slo_plot1 <- ggplot(pv_u_slo_plot) +
    theme_minimal() +
    geom_bar(aes(x = year, y = n_jobs, fill = occupation), size = 1.25, position = "stack", stat="identity")+
    scale_fill_manual(values = c("Construction" = '#4a4e69', "O&M" = '#9a8c98'))+ 
    #scale_x_continuous(expand = c(0, 0), breaks = scales::pretty_breaks()) +  # Remove space on x-axis
 # scale_y_continuous(expand = c(0, 0), breaks = scales::pretty_breaks()) +   # Remove space on y-axis 
    theme(plot.background = element_rect(color = "black", fill = NA, size = 1),
          #legend.text = element_blank(),
            axis.text.x = element_text(size = 16),  # Rotate x-axis labels
    axis.text.y = element_text(size = 16))+
    labs(x = NULL,
         y= "FTE jobs",
         title = "Projected direct jobs in San Luis Obispo county from Utility PV development")


pv_u_slo_plot1


ggsave('plots/pv_u_slo_plot.png', pv_all_plot1, width = 8, height = 6)
```

# OSW model

## High ambition

For the tri-county central coast JEDI run, we have:

Construction:
- direct: 1452
- indirect: 4313
- induced: 1179

O&M:
- direct: 140
- indirect: 112
- induced: 136

All in FTE jobs per GIGAWATT (JEDI would not let me run for 1 MW of OSW)

Remember that construction jobs will need to be / 5, and o&m jobs are annual.


Starting with direct construction jobs:
```{r}
# Create data frame, just adding year for now ----
year = 2025:2045
cap_table <- data.frame(
    year = year
)

# Ramping up total capacity from 3GW in 2030 to 15GW in 2045 ----
# Initial values
initial_capacity <- 3
final_capacity <- 15
years <- 15

# Calculate the annual growth rate ----
growth_rate <- (final_capacity / initial_capacity)^(1 / years) - 1

# Add total capacity for each year into empty list ----
total_capacity <- numeric(length(year))

for (i in 6:length(cap_table$year)) {
  total_capacity[i] <- initial_capacity * (1 + growth_rate)^(year[i] - 2030)
}  
  
# Add total capacity into df ----
cap_table$total_capacity <- total_capacity

# Add new capacity by the year construction starts ----
cap_table <- cap_table %>%
  mutate(
    new_construction_capacity = lead(total_capacity, 4) - lead(total_capacity, 3)
  )

# Fill NAs with 0 ----
cap_table$new_construction_capacity[is.na(cap_table$new_construction_capacity)] <- 0

# Add direct construction jobs ----
osw <- cap_table
osw$occupation = "construction"
osw$type = "direct"

# Initialize jobs/per GW Year that will be multiplied by capacity each year
jobs_gw_year <- 1452/5 # Total construction jobs outputted by JEDI / 5 Years
# New jobs each year
osw$new_jobs <- jobs_gw_year * osw$new_construction_capacity

# Total jobs ---- 
osw$total_jobs <- zoo::rollapply(osw$new_jobs, width = 5, FUN = sum, align = "right", partial = TRUE)
```

```{r}

# Initialize jobs/per GW Year that will be multiplied by capacity each year
variable_jobs <- 10 # Total construction jobs outputted by JEDI / 5 Years

# Create a tibble with years from 2026 to 2045
new_capacity <- 1:(2045-2026+1)  # Increment by 1 per year
new_jobs <- numeric(length(new_capacity))

total_jobs <- 0

for (i in seq_along(new_capacity)) {
        new_jobs[i] <- (new_capacity[i] * variable_jobs)
}
    

for (i in seq_along(new_capacity)) {
    if (i == 1) {
        total_jobs[i] <- new_jobs[i]  # find new jobs at year 1
    }else if (i <= 5) {
        total_jobs[i] <- total_jobs[i-1] + (new_capacity[i] * variable_jobs)
       # total_jobs[i] <- (new_capacity[i] * variable_jobs) + sum(new_jobs[1:i])  # Accumulate until year 6
    }else {
        total_jobs[i] <- (new_capacity[i] * variable_jobs) + sum(new_jobs[(i-4):(i-1)])  # Add current and previous 4 years' new jobs
    }
}

osw_df <- tibble(
    Year = 2026:2045,
    `New Jobs This Year` = new_jobs,
    `New Capacity` = new_capacity,
    `Total Jobs` = total_jobs  # Placeholder for future data entry
) 

print(osw_df)

ggplot(osw) +
    geom_line(aes(x = year, y = total_capacity)) +
    geom_line(aes(x = year, y = new_construction_capacity))
```


```{r}

ggplot(osw_df, aes(x = Year, y = new_jobs)) +
    geom_point() +
    labs(x = 'Year',
         y = 'Construction Jobs',
         title = 'High Ambition Offshore Wind Energy Targets') +
    theme_minimal()

ggplot(osw) +
    geom_col(aes(x = year, y = total_jobs)) +
    labs(title = "Projected direct construction jobs in CCC from OSW development",
         y = "FTE Jobs") +
    scale_y_continuous(labels = scales::label_comma()) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank()
    )

```

SWEET! Now we have direct construction jobs each year from 2025 - 2045. We can reuse this code to get indirect and induced jobs. 

Now, I want to try and do O&M jobs. This will be easier, since O&M is just multiplied straight across with total capacity

```{r}
# Add direct o&m jobs ----
osw_om_direct <- cap_table
osw_om_direct$occupation = "operations and maintenance"
osw_om_direct$type = "direct"

# Initialize jobs/per GW Year that will be multiplied total capacity
om_jobs_gw_year <- 140 # 140 annual o&m jobs / GW

# New jobs
osw_om_direct$new_jobs <- NA

# Total jobs
osw_om_direct$total_jobs <- om_jobs_gw_year * osw_om_direct$total_capacity

# combine into main osw df
osw <- rbind(osw, osw_om_direct)
```

```{r}
ggplot(osw) +
    geom_col(aes(x = year, y = total_jobs, fill = occupation),
             position = "stack") +
    labs(title = "Projected direct jobs in CCC from OSW development",
         y = "FTE Jobs") +
    scale_y_continuous(labels = scales::label_comma()) +
    scale_fill_manual(values = c("#4a4e69", "#9a8c98")) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank()
    )
```

Horray! In the future, we can go back and add indirect and induced for both occupations. Also maybe add a new row for ambition high or low and include both in one df?


Oil phaseout modeling 
We're going to start by reading in MSA_ALL_YEARS.csv. This is combined BLS data for fossil fuel and green job occupations.

```{r}
# Filter down data to our counties: Ventura, Santa Barbara, and San Luis Obispo
msa_all_years <- read.csv(here('MSA_ALL_YEARS.csv')) %>% 
   clean_names() %>% 
    filter(str_detect(area_title,
                      regex("Santa Barbara|Ventura|San Luis Obispo",
                            ignore_case = TRUE)))

# Filter only fossil fuel jobs in these counties
msa_ff <- msa_all_years %>% 
    select(-8:-22) %>% 
    filter(label == "ff job")

# Replace '**' with NA and then convert 'tot_emp' to numeric
msa_ff$tot_emp <- as.numeric(gsub("\\*\\*", NA, msa_ff$tot_emp))

# Find total ff jobs in each county (we'll use this to multiply by for scaledown)
msa_ff_grouped <- msa_ff %>% 
    group_by(area_title) %>% 
    summarize(total_employed = sum(tot_emp, na.rm = TRUE))
```

Now, let's plot our initial values for fossil fuel jobs in our respective areas. 
```{r}
ff_jobs_plot <- ggplot(msa_ff_grouped) +
    theme_minimal() +
    geom_col(aes(x = area_title, y = total_employed)) +
    theme(plot.background = element_rect(color = "black", fill = NA, size = 1),
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = NULL,
         y= "FF Jobs",
         title = "Number of Fossil Fuel Jobs in CC")

#ggsave('plots/ff_jobs_plot.png', ff_jobs_plot, width = 8, height = 6)
```

```{r}
# Create a new column for county grouping
msa_ff_final <- msa_ff_grouped %>% 
    mutate(county = case_when(
        str_detect(area_title, "Ventura") ~ "Ventura",
        str_detect(area_title, "Santa Barbara") ~ "Santa Barbara",
        str_detect(area_title, "San Luis Obispo") ~ "San Luis Obispo"
    ))

msa_ff_final %>% 
    group_by(county)

ff_jobs_plot_final <- ggplot(msa_ff_final) +
    geom_col(aes(x = county, y = total_employed)) +
    theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, margin = margin(5,20,0,10)),
        axis.text = element_text(size = 20), 
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        legend.position = "bottom",
        plot.background = element_rect(fill = "#EFEFEF"),
        plot.title = element_blank(),
        panel.grid = element_line(color = "grey85")
    ) +
    labs(x = NULL,
         y = NULL,
         title = "Number of Fossil Fuel Jobs in CC")

ggsave('plots/ff_jobs_plot.png', ff_jobs_plot_final, width = 8, height = 6)
```

