---
title: "Modeling Jobs"
format: html
warning: false
message: false
---

```{r}
library(tidyverse)
```


# Objective:

End goal is to have one data frame that looks something like this:

| county  | year | technology | new_capacity | total_capacity | occupation   | type    | n_jobs |
|---------|------|------------|--------------|----------------|--------------|---------|--------|
|   SLO   | 2026 | Utility PV | 1            | 1              | contruction  |direct   | 5      |
| ...     | ...  |  ...       | ...          | ...            | ...          |  ...    | ...    |
| Ventura | 2026 | OSW        | 1            | 1              | op and maint |indirect | 100    |

etc.

I imagine this coming together (concat) from many different dataframes: one for each county and technology

# PV model

Start with 1 piece and see if we can build this into a function from there

## Santa Barbara Utility Solar

Values from Eleanor's Tech. Scenarios sheet:

- Ambition: 722 MW by 2045
- Starting: 111 MW in 2025

First, use compounding interest to calculate yearly capacities:

```{r}
# Initial values
initial_capacity <- 111
final_capacity <- 722
years <- 20

# Calculate the annual growth rate
growth_rate <- (final_capacity / initial_capacity)^(1 / years) - 1

# Create variables to store the results
year <- 2025:2045
capacity <- numeric(length(year))
new_capacity <- numeric(length(year))

# Calculate the total capacity for each year
for (i in 1:length(year)) {
  capacity[i] <- initial_capacity * (1 + growth_rate)^(year[i] - 2025)
  if (i == 1) {
      new_capacity[i] <- capacity[i] - initial_capacity
  } else {
      new_capacity[i] <- capacity[i] - capacity[i-1]
  }
}

# Create a data frame with the results
sb_utility_pv <- data.frame(county = "SB", 
                             year = year, 
                             technology = "utility solar",
                             new_capacity = new_capacity,
                             total_capacity = round(capacity, 2))

# Display the table
print(sb_utility_pv)
```

Now we just need to add number of jobs! Let's just say for testing purposes that utility solar requires:
- 5 FTE construction jobs / MW
- construction project lasts 3 months
- 10 annual O&M jobs / MW
- lifetime of project is 30 years (beyond the scope of our model)  


Starting with O&M jobs:
```{r}
# O&M direct jobs ---- 
# Assign variables
om_direct <- 10

# Add columns occupation, type, and n_jobs
sb_utility_pv_om_direct <- sb_utility_pv |>
    mutate(occupation = "O&M",
           type = "direct",
           n_jobs = total_capacity * om_direct)

# O&M indirect jobs ---- 
# Assign variables
om_indirect <- 5

# Add columns occupation, type, and n_jobs
sb_utility_pv_om_indirect <- sb_utility_pv |>
    mutate(occupation = "O&M",
           type = "indirect",
           n_jobs = total_capacity * om_indirect)

# O&M induced ----
# Assign variables
om_induced <- 2

# Add columns occupation, type, and n_jobs
sb_utility_pv_om_induced <- sb_utility_pv |>
    mutate(occupation = "O&M",
           type = "induced",
           n_jobs = total_capacity * om_induced)

# Stack 3 together for total SB PV O&M jobs
sb_utility_pv_om <- rbind(sb_utility_pv_om_direct, sb_utility_pv_om_indirect, sb_utility_pv_om_induced)

print(sb_utility_pv_om)

```












# OSW model
